---
title: "Probability Cheatsheet"
author: "Felix Schneider"
date: "2025-02-07"
output: html_document
---

```{r include = F}
knitr::opts_chunk$set(error = T)
```

```{r include = F}
library(tidyverse)
```

## Propability of outcomes from casting two dice

### Creating samplespace

```{r}
omega <- expand.grid(x = 1:6, y = 1:6)
head(omega)
```

### Logical operations

creating subsets

```{r}
a <- omega %>% filter(x == 1)
a
b <- omega %>% filter(x + y == 7)
b
```

a and b

```{r}
intersect(a,b)
```

a or b

```{r}
union(a,b)
```

a and not b

```{r}
setdiff(a,b)
```

not a

```{r}
setdiff(omega,a) %>% head(10)
```

either a or b

```{r}
symdiff(a,b)
```

## Distributions

#### Example: we have 1000 students, 300 are female and 700 are male

```{r}
n <- 1000
p.f <- 0.3
n.f <- 300
p.m <- 0.7
n.m <- 700
```


### Binomial (with replacement)

#### dbinom
###### gives the probability that [30] students of [100] are female

```{r}
dbinom(x = 30,
       size = 100,
       prob = p.f)
```

###### also used to graph probability distribution

```{r}
plot(
  x = 0:100,
  y = dbinom(
    x = 0:100,
    size = 100,
    prob = p.f
    ),
  type = "l",
  xlab = "n",
  ylab = "prob"
)
```

#### pbinom
###### gives the cumulative probability that a maximum of [30] students out of [100] are female

```{r}
pbinom(q = 30,
       size = 100,
       prob = p.f)
```

###### also used to plot cumulative probability distribution

```{r}
plot(
  x = 0:100,
  y = pbinom(
    q = 0:100,
    size = 100,
    prob = p.f
    ),
  type = "l",
  xlab = "n",
  ylab = "prob"
)
```

#### qbinom
###### finds the cutoff where a given probability is reached in a binomial distribution use it when you need thresholds for passing, ranking, or decision-making

```{r}
qbinom(
  p = 0.5,
  size = 100,
  prob = p.f
)
```

###### graph not strictly neccessary but for visual aid

```{r}
plot(
  x = seq(0,1,0.001),
  y = qbinom(
    p = seq(0,1,0.001),
    size = 100,
    prob = p.f
    ),
  type = "l",
  xlab = "prob",
  ylab = "n"
)
```

#### rbinom
###### simulates [10] random samples of [100] people and returns the number of female students in ea

```{r}
rbinom(10,100,p.f)
```

### Hypergeometric (Without replacement)

#### dhyper
###### gives the probability that [30] students of [100] are female

```{r}
dhyper(x = 30,
       m = n.f,
       n = n.m,
       k = 100)
```

###### also used to graph probability distribution

```{r}
plot(
  x = 0:100,
  y = dhyper(
    x = 0:100,
    m = n.f,
    n = n.m,
    k = 100
    ),
  type = "l",
  xlab = "n",
  ylab = "prob"
)
```

#### phyper
###### gives the cumulative probability that a maximum of [30] students out of [100] are female

```{r}
phyper(q = 30,
       m = n.f,
       n = n.m,
       k = 100)
```

###### also used to plot cumulative probability distribution

```{r}
plot(
  x = 0:100,
  y = phyper(
    q = 0:100,
    m = n.f,
    n = n.m,
    k = 100
    ),
  type = "l",
  xlab = "n",
  ylab = "prob"
)
```

#### qhyper
###### finds the cutoff where a given probability is reached in a binomial distribution use it when you need thresholds for passing, ranking, or decision-making

```{r}
qhyper(
  p = 0.5,
  m = n.f,
  n = n.m,
  k = 100
)
```

###### graph not strictly neccessary but for visual aid

```{r}
plot(
  x = seq(0,1,0.001),
  y = qhyper(
    p = seq(0,1,0.001),
    m = n.f,
    n = n.m,
    k = 100,
    ),
  type = "l",
  xlab = "prob",
  ylab = "n"
)
```

#### rhyper
###### simulates [10] random samples of [100] people and returns the number of female students in ea

```{r}
rhyper(nn = 10,
        m = n.f,
        n = n.m,
        k = 100)
```












