---
title: "Plots Cheatsheet"
author: "Felix Schneider"
date: "2025-02-06"
output: html_document
---

```{r include = F}
knitr::opts_chunk$set(error = T)
```

```{r include = F}
library(tidyverse)
```


```{r}
election <- tibble(party = c("CDU","SPD","AFD","FDP","DIE LINKE","GRUENE","CSU","OTHER"),
                   result.13 = c(26.8,20.5,12.6,10.7,9.2,8.9,6.2,5.0),
                   result.17 = c(34.1,25.7,4.7,4.8,8.6,8.4,7.4,6.2)) %>% 
  mutate(difference = result.17 - result.13)

election
```


## Barplot

```{r}
election$result.17 %>% barplot(names.arg = election$party,
                               ylim = c(0,50),
                               xlab = "Parties",
                               ylab = "2017 Vote %",
                               las = 2, #rotates names
                               beside = T,
                               col = rainbow(8),)

election$difference %>% barplot(names.arg = election$party,
                                ylim = c(-10,10),
                                xlab = "Parties",
                                ylab = "Vote difference 13-17",
                                las = 2,
                                beside = T,
                                col = rainbow(8)
                                )

rbind(election$result.13,election$result.17) %>%
  barplot(names.arg = election$party,
          ylim = c(0,50),
          xlab = "Parties",
          ylab = "Vote %",
          las = 2,
          beside = T,
          col = c("green","purple"),
          legend.text = c("2013","2017")
          )
```

## Piechart

```{r}
election$result.17 %>% pie(labels = election$party,
                           col = rainbow(8),
                           main = "Election 2017")
```

## Histogram

```{r}
times <- c(568, 577, 577, 581, 640, 641, 645, 657, 673, 696, 703, 720,
           728, 728, 728, 729, 777, 808, 824, 825, 865, 975,1007,1007)

hist(times)

bounds <- c(500,700,950,1100)
hist(times,
     breaks = bounds,
     xlab = "time",
     col = rainbow(3))
```

## Cumulative Frequency Distribution

```{r}
plot.ecdf(times,
          main = "cumulative Frequency")

bounds <- c(500,600,700,800,900,1000,1100)
times.cut <- cut(
  times,
  breaks = bounds,
  labels = bounds [-1]
  )
plot.ecdf(
  as.integer(as.character(times.cut)),
  xlab = "times", ylab = "H(x)",
  main = "H Based on grouped Data"
  )
```

get the amount of results <= 800

```{r}
H <- ecdf(times)
H(800)
```

## Boxplot

```{r}
df <- tibble(
  non_player = c(22.1, 22.3, 26.2, 29.6, 31.7,
                 33.5, 38.9, 39.7, 43.2, 43.2),
  beginner   = c(32.5, 37.1, 39.1, 40.5, 45.5,
                 51.3, 52.6, 55.7, 55.9, 57.7),
  tournament = c(40.1, 45.6, 51.2, 56.4, 58.1,
                 71.1, 74.9, 75.9, 80.3, 85.3)
  )

head(df)

df %>%
  pivot_longer(
    cols = everything(),
    names_to = "group"
  ) %>% 
  group_by(group) %>% 
  summarise(
    mean = mean(value),
    min = min(value),
    Q1 = quantile(value,0.25,1),
    median = median(value),
    Q3 = quantile(value,0.75,1),
    max = max(value),
    IQR = quantile(value,0.75,1)-quantile(value,0.25,1),
    variance = var(value)
  )

df %>% boxplot()

MPG <- mpg %>% select(displ,hwy)%>%
  mutate(displ_class = case_when(displ > 1 & displ <= 3 ~ "low",
                                 displ > 3 & displ <= 5 ~ "medium",
                                 displ > 5 & displ <= 8 ~ "high"))
head(MPG)

boxplot(MPG$hwy ~ MPG$displ_class,
        xlab = "Displacement",
        ylab = "Highway miles per galon",
        col = rainbow(3))
```

## Scatterplot and linear regression

```{r}
df <- tibble(
  x = c(2, 6, 3, 4, 5),
  y = c(3, 7, 4, 7, 6)
  )

plot(
  x = df$x,
  y = df$y,
  col = "blue"
) #this part is not executable line by line if there is no plus here
  # but it will still work when executed as a block or knited
abline(lm(df$y ~ df$x),
       col = "red")

line <- lm(df$y ~ df$x)
a <- line$coefficients[1]
b <- line$coefficients[2]
c(a,b)

corelation = cor(x = df$x, y = df$y)
corelation

covariance = cov(x = df$x, y = df$y)
covariance
```


















