---
title: "Tibbles and Tidy Cheatsheet"
author: "Felix Schneider"
output: html_document
date: "`r Sys.Date()`"
---

```{r include = F}
knitr::opts_chunk$set(error = T)
```

```{r include = F}
library(tidyverse)
```

## Pivot_longer

Simple case

```{r}
df <- tibble(
  student = c("Adam","Bernd"),
  analysis = c(3,1),
  algebra = c(4,3)
)

df

df %>% 
  pivot_longer(
    cols = !student,
    names_to = "subject",
    values_to = "grade"
  )
```

More complex case where columns have common prefixes and NAs should be dropped

```{r}
df <- tibble(
  subject = c("algebra", "analysis"),
  wk1 = c(5,8),
  wk2 = c(7,NA)
)

df

df %>% 
  pivot_longer(
  cols = starts_with("wk"),
  names_to = "week",
  names_prefix = "wk",
  values_to = "attendees",
  values_drop_na = T
)
```

Multiple Variables stored in column names

```{r}
df <- tibble(
  country = "germany",
  new_a0_m = 12,
  new_a0_f = 23,
  newb2_m = 43,
  newb2_f = 50,
)

df

df %>% 
  pivot_longer(
    cols = new_a0_m:newb2_f,
    names_to = c("type", "gender"),
    names_pattern = "new_?(.*)_(.)",
    values_to = "count"
  )
```

## Pivot_wider

Siple case

```{r}
df <- tibble(
  subject = c("analysis","algebra","analysis"),
  student = c("Bernd","Bernd","Adam"),
  points = c(10,5,10)
)

df

df %>% 
  pivot_wider(
    names_from = subject,
    values_from = points
  )
```

Filling a standard value where there is an NA

```{r}
df %>% 
  pivot_wider(
  names_from = subject,
  values_from = points,
  values_fill = 0
)
```

Generate column names from multiple variables

```{r}
df <- tibble(
  name = rep(c("Adam", "Bernd"),each=2),
  variable = rep(c("rent", "income"), 2),
  estimate = c(413,746,823,251),
  moe = c(837,376,262,436)
)

df

df %>%  
  pivot_wider(
    names_from = variable,
    values_from = c(estimate,moe)
  )
```

changing the way the new variables look

```{r}
df %>%  pivot_wider(
  names_from = variable,
  values_from = c(estimate,moe),
  names_sep = ":"
)
```

executing a function in the pivot

```{r}
df <- tibble(
  breaks = sample(20:40,12),
  wool = rep(c("A","B"),each=6),
  tension = rep(rep(c("L","M","H"),each = 2),2)
)

df

df %>% pivot_wider(
  names_from = wool,
  values_from = breaks,
  values_fn = mean
)
```



## Separate

simple example

```{r}
df <- tibble(
  name = c("Adam", "Bernd", "Christian", "Doris"),
  ratio = c("81/1.83", "71/1.75", "55/1.69", "62/1.57")
  )

df

df %>%  
  separate(
    col = "ratio",
    into = c("weight", "height"),
    sep = "/"
  ) 
```

delete separated part by calling it NA

```{r}
df %>%  
  separate(
    col = "ratio",
    into = c(NA, "height"),
    sep = "/"
  )  
```



## Tibble interaction

```{r}
df <- tibble(
  sex = c("f", "m", "f", "f", "m"),
  age = sample(20:35, 5, replace = T),
  score1 = round(runif(5, 0, 25))
)

df
```

### add_row

```{r}
df2 <- df %>% 
  add_row(
    sex = "m",
    age = 25,
    score1 = 20)

df2
```

### mutate and case_when

```{r}
df3 <- df2 %>% mutate(
  score2 = sample(0:25,6,replace = T),
  score3 = round(runif(6, 0, 25)),
  score.total = score1 + score2 + score3,
  grade = case_when(
     score.total <= 37 ~ 5,
     score.total > 37 & score.total <= 45 ~ 4,
     score.total > 45 & score.total <= 55 ~ 3,
     score.total > 55 & score.total <= 65 ~ 2,
     score.total > 65 ~ 1
  )
)

df3
```

### filter

```{r}
df3 %>% filter(sex == "m")
```

### select

```{r}
df3 %>% select(
  sex, score.total, grade
)

df3 %>% select(score1:score3)
```

### group_by and summarise

```{r}
df3 %>% 
  group_by(sex) %>% 
  summarise(
    count = n(),
    avrg.score = mean(score.total),
    avrg.grade = mean(grade)
    )

df3 %>% 
  group_by(sex, grade) %>% 
  count()
```

### Arrange

```{r}
df3 %>% arrange(score.total)

df3 %>% arrange(-score.total)

df4 <- df3 %>% 
  arrange(sex,score.total)

df4
```



### Head and Tail

```{r}
df4 %>% head(3)

df4 %>% tail(3)
```



























